---
title: "Report_Chapter4"
author: "Chiara Huwiler"
date: "2023-03-20"
output: html_document
---

## AGDS 4.4 Report Exercises

# Telling a story from data

This dataset contains the daily air quality measurements in New York from May to September 1973. Obtained is the data from the New York State Department of Conversation an the National Weather Service. With three relevant patterns the hypothesis: *"How is the solar radiation interfering with the ozone?"* is analysed.

```{r}
# Setting up
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r}
# loading and defining the Dataset
datasets::airquality
airquality1 <- datasets::airquality

#renaming of the colume Month (to be better distuinguish, from the function month)
airquality1 <- airquality1 |>
  rename(Month_1 = Month)
knitr::kable(airquality1)
```

# Patterns

## Identifying Outliers

First the outliers were identified. That means that extreme values which could have a significant impact on the analysis or modeling of the data are identified.

```{r}
# Computing the outliers
vec_outliers <- boxplot.stats(airquality1$Ozone)$out

# Creating new data frame including variable outlier
plot_o <- airquality1 |> 
  mutate(outlier = Ozone %in% vec_outliers)

# Creating a scatterplot with marked outliers

plot_o1<- plot_o |> 
  ggplot(aes(x = Solar.R, y = Ozone, color = outlier)) + 
  geom_point() + 
  scale_color_manual("Outlier",                    # Sets title of legend
                     values = c("black", "red"),    # Highlights in red
                     labels = c("No", "Yes")        # Adds labels to the legend
                     ) +
  labs(title = "Outliers", subtitle = "May to September 1973",
    x = expression(paste("Solar (in the frequency band 4000-7700)")), 
       y = expression(paste("Ozone (in parts by billion)"))) +
  theme_classic()

# Viewing the plot
print(plot_o1)
```

```{r}
# Saving this plot
ggsave("./figures/plot_o.jpeg", plot = plot_o1, device = "jpeg")
```

The plot shows that there is a general positive correlation between solar radiation and mean ozone level, as higher solar radiation levels tend to be associated with higher mean ozone levels. Here, two outlines are identified. About them it needs to be taken extra caution to get good data.

## Scatterplot

Secondly a scatterplot is used. It visualize how two variables co-vary and how an added third variable (here in b, c and d) interferes with them.

```{r}
# preparing plot data
airquality1_subset <- airquality1

# a: Regressing Solar against Ozone
plot_a <- ggplot(
  data = airquality1_subset,
  aes(x = Solar.R, y = Ozone)) +
  geom_point(size = 0.75) +
  geom_smooth(method = "lm", color = "red") +
  labs(x = expression(paste("Solar.R (in the frequency band 4000--7700)")), 
       y = expression(paste("Ozone (per Billion)"))) +
  theme_classic()

# b: Mapping the variable Temp on the plot
plot_b <- ggplot(
  data = airquality1_subset,
  aes(x = Solar.R, y = Ozone, color = Temp)) +
  geom_point(size = 0.75) +
  labs(x = expression(paste("Solar.R (in the frequency band 4000--7700)")), 
       y = expression(paste("Ozone (per Billion)"))) +
  theme_classic()

# c: Mapping Months as a continous variable on the plot
plot_c <- ggplot(
  data = airquality1_subset,
  aes(x = Solar.R, y = Ozone, color = as.factor(Month_1))) +
  geom_point(size = 0.75) +
  labs(x = expression(paste("Solar.R (in the frequency band 4000--7700)")), 
       y = expression(paste("Ozone (per Billion)"))) +
  theme_classic()

# d: Mapping Wind on the plot
plot_d <- ggplot(
  data = airquality1_subset,
  aes(x = Solar.R, y = Ozone, color = Wind)) +
  geom_point(size = 0.75) +
  labs(x = expression(paste("Solar.R (in the frequency band 4000--7700)")),
       y = expression(paste("Ozone (per Billion)"))) +
  theme_classic() +
  scale_color_viridis_c()

# combine plots
plota_d <- cowplot::plot_grid(plot_a, plot_b, plot_c, plot_d, ncol = 2, labels = "auto") 

# Viewing the plots
print(plota_d)
```

```{r}
# Saving this plot 
ggsave("./figures/plota_d.jpeg", plot = plota_d)

```

Here the results tell that a higher Temperature, as well as less wind leads to an higher ozone level. Whereas within the months no trend is recognizable.

## Regression within categories

Lastly the regression within categories is plotted. Here the months (as categories) are used for separating the data, so each color represents a month.

```{r}
# Converting the Month_1 variable in a factor
airquality1$Month_1 <- factor(airquality1$Month_1)

# Creating scatterplot
plot_r <- ggplot(
  data = airquality1,
  aes(x = Solar.R, y = Ozone, color = Month_1)) +
  geom_point(alpha = 0.5) + # sets transpearency
  geom_smooth(formula = y ~ x + 0, method = "lm", se = FALSE) + # adds a liner regression
  labs(x = expression(paste("Solar")), 
       y = expression(paste("Ozone")) ) +
  theme_classic() +  scico::scale_color_scico_d(palette = "romaO")

# Viewing the plot
print(plot_r)
```

```{r}
# Saving this plot
ggsave("./figures/plot_r.jpeg", plot = plot_r)
```

Here it is suggested that solar radiation has a positive linear relationship with ozone levels. But within this relationship the values are different depending on the month. In June and July is the ozone level in tendency higher.

# Interpretation and discussion

The above generated plots are providing data about how the solar radiation is interfering with the ozone level and so with the air quality in New York City. A positive relationship can be seen between the solar radiation and the ozone level. The higher the radiation the higher the ozone level seems to be. Additionally we can see that the temperature affect this relationship (see regression plot), as well as the trend of ozone level changes during the year (see scatterplot), e. p. a higher ozone level in June and July.
