---
title: "Report_Chapter3"
author: "Chiara Huwiler"
date: "2023-03-06"
output: html_document
---

```{r}
# Setting up
library(tidyverse)
```

## AGDS 3.5 Report Exercises

This document includes the solutions to all of the Chapter 3 Exercises from the Applied Geodata Science course.

This exercises are based on the following database: Groenigen, Kees Jan van, Xuan Qi, Craig W. Osenberg, Yiqi Luo, and Bruce A. Hungate. "Faster Decomposition Under Increased Atmospheric CO2 Limits Soil Carbon Storage." Science 344, no. 6183 (May 2, 2014): 508--9. <https://doi.org/10.1126/science.1249534>.

The data is imported as a tibble

```{r}
# Reading and defining the dataframe 
df_full <- readr::read_csv2("./df_full.csv", col_names = TRUE)

```

# Excercise 2

Data aggregation per experiment and calculation of the log-response ratio

RR = ln (xelevated/xambient)

For that calculate for each Experiment the means for each Column.

```{r}
# Calculation the mean values
df_exp <- df_full |>
  group_by(Experiment) |>
  summarise (ambient_mean = mean(ambient_mean),
            elevate_mean = mean(elevate_mean))

# Displaying the result
knitr::kable(df_exp)

```

Calculate the log response ratio of every Experiment from the new Database mean_eachexperiment

```{r}
# Calculation the log response ratio
df_lrr <- df_exp |>
  reframe ( LRR <- log(elevate_mean/ambient_mean))


# Displaying the result
knitr::kable(df_lrr)
```

# Exercise 3

Data aggregation across all experiments for different years (early, mid- and late phase); Calculation of the log-response ratio for each phase.

**Early Phase**

```{r}
# Defining the time period
early_phase <- df_full [df_full[,"Time"] >= 0 & df_full[,"Time"]<= 3,] |>
mutate(Time = as.factor(Time))  |>

# Calculating means of ambient_mean and elevate_mean
summarise(ambient_mean = mean(ambient_mean),
            elevate_mean = mean(elevate_mean)) |>

# Calculating log-response ratio
mutate (ep_mean <- log(elevate_mean/ambient_mean))

# Displaying the result
knitr::kable(early_phase)
```

**Midphase**

```{r}
# Defining the time period
midphase <- df_full [df_full[,"Time"] >= 3 & df_full[,"Time"]<= 6,] |>
mutate(Time = as.factor(Time))  |>

# Calculating means of ambient_mean and elevate_mean
summarise(ambient_mean = mean(ambient_mean),
            elevate_mean = mean(elevate_mean)) |>

# Calculating log-response ratio
mutate (ep_mean <- log(elevate_mean/ambient_mean))

# Displaying the result
knitr::kable(midphase)
```

**Late phase**

```{r}
# Defining the time period
late_phase <- df_full [df_full[,"Time"] >= 6,] |>
mutate(Time = as.factor(Time))  |>

# Calculating means of ambient_mean and elevate_mean
summarise(ambient_mean = mean(ambient_mean),
            elevate_mean = mean(elevate_mean)) |>

# Calculating log-response ratio
mutate (ep_mean <- log(elevate_mean/ambient_mean))

# Displaying the result
knitr::kable(late_phase)
```
